---
title: "R Notebook"
output: html_notebook
---

```{r}
library(parallel)
library(data.table)
library(SANjar)
library(SANjar.data.lt48)
library(ggplot2)
```

Define posterior distribution
=============================

## First, define a parametrization

Here, we define a single variable that gets used for $g_S$, $g_A$ up to day 40. We further specify that deviations of the rank-size distribution should be computed only on days 11 and 40, and that S-cell counts should not be reported for any day.

```{r}
pr <- san_parametrization(list(`40SA`=range(0, 4)), basemodel=lt47.model,
                          rs_days=c(11, 40), sc_day=c())
```

## Then, turn the parametrization into a posterior likelihood

The data to evaluate the model against must be provided as an object of type "LTData". Here, we just use the pre-packages LT47 data.

```{r}
pd <- san_posterior(pr, lt47)
```

## Evaluating posterior likelihoods

Using the `loglikelihood` function in the posterior distribution object, the posterior likelihood can be evaluated for a single or multiple combinations of parameters.

The result contains the following columns

| Field            | Description |
|------------------|----------------------------------|
| `ll_tot`         | the total posterior likelihood   |
| `ll_cc`          | the partial posterior likelihood for total cell counts |
| `ll_rs_<day>`    | the partial poisterior likelihood for rank-sizes on day `<day>` |
| `CC<day>`        | the total cell count on day `<day>` |
| `SC<day>`        | the S-cell count on day `<day>` |
| `RS<day>_<rank>` | the rank-size for rank `<rank>` on day `<day>` |

```{r}
ll <- pd$loglikelihood(data.table(`40SA`=c(1.4,1.5,1.6,1.7)))
ll
```

The resulting trajectories can be plotted against the empirically observed intervals

```{r}
plot(pd, ll, rs_days=c(11, 40))
```

## Combining posterior distributions

Two posterior distribution objects can be combined into a single object, which results in a distribution whose likelihood is the sum of the two likelihoods. This can be used to evaluate parameters for two separate datasets and combine the results

```{r}
pr.lt48 <- san_parametrization(list(`40SA`=range(0, 4)), basemodel=lt47.model,
                               rs_days=c(11, 42), sc_day=c())

pd.wt <- san_posterior(pr.lt48, lt48.knockout[condition="PAX6", lib="wt"])
pd.ko <- san_posterior(pr.lt48, lt48.knockout[condition="PAX6", lib="ko"])
pd.wt_and_ko <- san_posterior_combine(wt=pd.wt, ko=pd.ko)
```

The result returned by `loglikelihood` then contains column `ll_tot` containing the total likelihood (i.e. the sum of the individual likelihoods of the component distributions) and one set of meta-data columns for each component, prefixed with the component name specified when the combined posterior distribution object was created.

```{r}
ll.wt_and_ko <- pd.wt_and_ko$loglikelihood(data.table(`40SA`=c(1.6,1.7)))
ll.wt_and_ko
```
```{r}
plot(pd.wt_and_ko, ll.wt_and_ko, rs_days=c(42))
```

## Parallelizing likelihood computations

First, a cluster must be created using one of the cluster creation functions from the `parallel` package

```{r}
cl <- makePSOCKcluster(8)
```

Then, a parallel version of an existing posterior distribution object can be created using `san_posterior_parallel`

```{r}
pd.parallel <- san_posterior_parallel(pd, cl)
```

Parallel distribution objects are used just as serial ones

```{r}
ll.par <- pd.parallel$loglikelihood(data.table(`40SA`=c(1.4,1.5,1.6,1.7)))
ll.par
```

The resulting trajectories can be plotted against the empirically observed intervals

```{r}
plot(pd, ll.par, rs_days=c(11, 40))
```

Running MCMC
============

The `mcmc` function samples from the specified posterior distribution. Each chain is run for the specified number of steps, and it's final value is used a one sample. The `candidates` are used to initialize the chains. If more candidates than chains are specified, the candidates are split across chains and are successively used as proposals in an independent-proposal MCMC to select each chain's starting value.

```{r}
mc <- mcmc(pd.parallel$loglikelihood, variables=pr$ranges, chains=8, steps=10,
           candidates=sample.variables(40, pr$ranges), verbose=TRUE)
```

The resulting obtains contains the final value of each chain, which represent (if the number of steps was sufficient) independent samples taken from the posterior,

```{r}
print(mc$final)
```

The posterior density can be plotted as violin plots

```{r}
plot(mc)
```

and various summary statistics of the posterior distribution can be computed

```{r}
summarystats(mc)
```
It's also possible to plot the individual trajectories for the final parameter values

```{r}
plot(pd, mc$final, rs_days=c(40))
```